---
title: 'Reproducible Research: Peer Assessment 1'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
```{r read data, echo=TRUE}
activity_data <- read.csv("activity.csv")
activity_data$date <- as.POSIXct(strptime(activity_data$date,"%Y-%m-%d"))
activity_data$date <- as.Date(activity_data$date)
library(tidyverse)

```


## What is mean total number of steps taken per day?
```{r total_steps,echo=TRUE}

total_steps <- aggregate(activity_data$steps,by = list(Category = activity_data$date),FUN = sum)

hist(total_steps$x,xlab = "Total Steps Per Day", ylab = "Frequency", main = "Histogram of total steps")

print(mean(total_steps$x,na.rm = TRUE))
print(median(total_steps$x,na.rm = TRUE))
```


## What is the average daily activity pattern?
```{r activity_pattern,echo=TRUE}
library(tidyverse)
complete_data <- na.omit(activity_data)
daily_steps <- tapply(complete_data$steps, complete_data$interval, mean)
intervals <- as.integer(names(daily_steps))
step_list <- data.frame("Steps" = daily_steps,"Interval" = intervals) 


steps_plot <- ggplot(step_list,aes(x = Interval,y = Steps)) +
        geom_line(col = "blue") 

print(steps_plot)

Max_steps <- filter(step_list,Steps == max(step_list$Steps))$Interval

```

### Maximum steps
Maximum steps is in the interval value of `r Max_steps`

## Imputing missing values

```{r}
missing_data <- is.na(activity_data)
sum_na_data <- sum(missing_data)


print(sum_na_data)

# Function for replacing NA step values from same interval means, 
data_fill <- function(data_set){
        for(i in seq_along(data_set[,1])){
                index = data_set[i,3]
                if(is.na(data_set[i,1])){data_set[i,1] <- mean(subset(data_set,interval==index)$steps,na.rm = TRUE)}
        }
return(data_set)
}

filled_data <- data_fill(activity_data)

filled_steps <- aggregate(filled_data$steps,by = list(Category = activity_data$date),FUN = sum)

hist(filled_steps$x,xlab = "Total Steps Per Day", ylab = "Frequency", main = "Histogram of total steps for filled data")

print(mean(filled_steps$x,na.rm = TRUE))
print(median(filled_steps$x,na.rm = TRUE))

```

### Mean and Median Values for Steps

```{r}
#Mean
print(mean(filled_steps$x,na.rm = TRUE))
#Median
print(median(filled_steps$x,na.rm = TRUE))
```

Filling missing data with interval means did not have much effect on the distribution of the data as the mean and median stayed practically the same.


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
library(lubridate)
activity_data <- mutate(activity_data, Days = wday(date))
activity_data <- mutate(activity_data, Weekend = as.integer(Days > 5))
activity_data$Weekend <- as.factor(activity_data$Weekend)

variable_names <- list("0" = "Weekdays","1" = "Weekends")
variable_labeller <- function(variable,value){
  return(variable_names[value])
}

mean_activity <- activity_data %>%
  group_by(Weekend,interval) %>%
  summarise(
    means_Steps = mean(steps,na.rm = TRUE)
  )

ggplot(mean_activity,aes(interval,means_Steps)) +
        geom_line(col= "blue") +
        facet_wrap(~Weekend,nrow = 2,ncol = 1,labeller =   variable_labeller) +
        xlab("Intervals") + 
        ylab("Mean Steps")
        



        
```

On general looks like weekends has higher activity than weekdays. 

Especially, the plots of weekday and weekend activity hints that at the weekends late day activity is more prominent compared to weekday late day activities. 


