---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data
```{r read data, echo=TRUE}
activity_data <- read.csv("../dataset/activity.csv")
activity_data$date <- as.POSIXct(strptime(activity_data$date,"%Y-%m-%d"))
activity_data$date <- as.Date(activity_data$date)

```


## What is mean total number of steps taken per day?
```{r mean_steps,echo=TRUE}
mean_steps <- mean(activity_data$steps,na.rm = TRUE)
print(mean_steps)
```

Mean total number of steps are `r mean_steps`

## What is the average daily activity pattern?
```{r activity_pattern,echo=TRUE}
library(tidyverse)
complete_data <- na.omit(activity_data)
daily_steps <- tapply(complete_data$steps, complete_data$interval, mean)
intervals <- as.integer(names(daily_steps))
step_list <- data.frame("Steps" = daily_steps,"Interval" = intervals) 


steps_plot <- ggplot(step_list,aes(x = Interval,y = Steps)) +
        geom_line(col = "blue") 

print(steps_plot)

Max_steps <- filter(step_list,Steps == max(step_list$Steps))$Interval

```

### Maximum steps
Maximum steps is in the interval value of `r Max_steps`

## Imputing missing values

```{r}
missing_data <- is.na(activity_data)
sum_na_data <- sum(missing_data)

# Function for completing NA step values from interval means
data_fill <- function(data_set){
        for(i in seq_along(data_set[,1])){
                index = data_set[i,3]
                if(is.na(data_set[i,1])){data_set[i,1] <- mean(subset(data_set,interval==index)$steps,na.rm = TRUE)}
        }
return(data_set)
}

filled_data <- data_fill(activity_data)

hist(filled_data$steps,xlab = "Total Steps", main = "Histogram of Total Steps per Day")

mean_fill <- mean(filled_data$steps)
median_fill <- median(filled_data$steps)
```

### Mean and Median Values for Steps
Mean steps for the completed activity data is `r mean_fill`.

Median steps for the completed activity data is `r median_fill`.



## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
library(lubridate)
activity_data <- mutate(activity_data, Days = wday(date))
activity_data <- mutate(activity_data, Weekend = as.integer(Days > 5))
activity_data$Weekend <- as.factor(activity_data$Weekend)

variable_names <- list("0" = "Weekdays","1" = "Weekends")
variable_labeller <- function(variable,value){
  return(variable_names[value])
}

mean_activity <- activity_data %>%
  group_by(Weekend,interval) %>%
  summarise(
    means_Steps = mean(steps,na.rm = TRUE)
  )

ggplot(mean_activity,aes(interval,means_Steps)) +
        geom_line(col= "blue") +
        facet_wrap(~Weekend,nrow = 2,ncol = 1,labeller =   variable_labeller) +
        xlab("Intervals") + 
        ylab("Mean Steps")
        



        
```


